---
title: "Introduction to Linear Models: 3 - Simple linear regression"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(emo)
library(learnr)
library(tidyverse)
library(openintro)

knitr::opts_chunk$set(echo = FALSE, 
                      fig.align = "center", 
                      fig.height = 3, 
                      fig.width = 5,
                      warning = FALSE, 
                      message = FALSE)

tutorial_options(exercise.eval = FALSE)
```

## Visualization of  Linear Models
### Possums

Before we get into the mathematical specification for a regression model, let's build some intuition about what a regression line is.  

In this scatterplot, we see the relationship between the total length of a possum, and the corresponding length of its tail. Clearly there is not a perfect relationship here: the total length of the possum varies even for possums with the same tail length. But we still have some intuitive desire to describe the relationship with a line. 

```{r 1, message=FALSE}
ggplot(data = possum, aes(y = total_l, x = tail_l)) +
  geom_point()
```

### Through the origin

This line goes through the origin - that is, the point where both $x$ and $y$ are equal to zero - and has a slope of 2.5 centimeters (of total length) per centimeter (of tail length). In some sense, it does go "through" the points, but doesn't capture the general trend as best we could imagine. 

```{r 2, message=FALSE}
ggplot(data = possum, aes(y = total_l, x = tail_l)) +
  geom_point() + geom_abline(intercept = 0, slope = 2.5)
```

### Through the origin, better fit

This line also goes through the origin, but has a gentler slope. It seems like a "better" fit, since it cuts through the points in a more central way.  

```{r 3, message=FALSE}
ggplot(data = possum, aes(y = total_l, x = tail_l)) +
  geom_point() + geom_abline(intercept = 0, slope = 2.3)
```

### Not through the origin

But why should we force the line to go through the origin? Here is a line that has a $y$-intercept of 40 cm, and an even gentler slope of 1.3 cm (of total length) per cm (of tail length). It seems like an even better fit still. 

Do you think you could find an even better fit? In order to do so, you need some criteria for judging which line fits better. In particular, you need a numerical measurement of how good the fit of each possible line is. 

```{r 4, message=FALSE}
ggplot(data = possum, aes(y = total_l, x = tail_l)) +
  geom_point() + geom_abline(intercept = 40, slope = 1.3)
```

### The "best" fit line

In regression, we use the least squares criterion to determine the best fit line. Statisticians have proven that (apart from pathological examples) if we seek the line that tries to minimize the sum of the squared distances between the line and a set of data points, a unique line exists. That line is called the least squares regression line. We can add the line to our plot using the `geom_smooth()` function and specifying the `method` argument to be `lm`, which stands for "linear model". 

```{r 5, message=FALSE}
ggplot(data = possum, aes(y = total_l, x = tail_l)) +
  geom_point() + geom_smooth(method = "lm")
```

### Ignore standard errors

Note that by default, this will draw the regression line in blue, with gray shading for the standard error associated with the line. That should not concern us just yet, so we can turn it off by setting the `se` argument to `FALSE` or to 0. 

```{r 6, message=FALSE}
ggplot(data = possum, aes(y = total_l, x = tail_l)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
```

You'll explore the "best fit" line on your own in these next exercises.

### The "best fit" line

The simple linear regression model for a numeric response as a function of a numeric explanatory variable can be visualized on the corresponding scatterplot by a straight line. This is a "best fit" line that cuts through the data in a way that minimizes the distance between the line and the data points. 

We might consider linear regression to be a specific example of a larger class of *smooth* models. The `geom_smooth()` function allows you to draw such models over a scatterplot of the data itself. This technique is known as visualizing the model *in the data space*. The `method` argument to `geom_smooth()` allows you to specify what class of smooth model you want to see. Since we are exploring linear models, we'll set this argument to the value `"lm"`.

Note that `geom_smooth()` also takes an `se` argument that controls the standard error, which we will ignore for now.


Create a scatterplot of body weight as a function of height for all individuals in the `bdims` dataset with a simple linear model plotted over the data.


```{r ex1, exercise = TRUE}
# Scatterplot with regression line
 ggplot(data = ___, aes(x = ___, y = ___)) + 
  ___ + 
  ___(method = ___, se = FALSE)
```

```{r ex1-hint}
 ggplot(data = ___, aes(x = ___, y = ___)) + 
  geom_point() + 
  ___(method = ___, se = FALSE)
```

```{r ex1-solution}
# Scatterplot with regression line
ggplot(data = bdims, aes(x = hgt, y = wgt)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)
```

### Uniqueness of least squares regression line

The least squares criterion implies that the slope of the regression line is unique. In practice, the slope is computed by R. In this exercise, you will experiment with trying to find the optimal value for the regression slope for weight as a function of height in the `bdims` dataset via trial-and-error.

To help, we've built a custom function for you called `add_line()`, which takes a single argument: the proposed slope coefficient.


The `bdims` dataset is available in your workspace. Experiment with different values (to the nearest integer) of the `my_slope` parameter until you find one that you think fits best.

```{r ex2-setup}
add_line <- function (my_slope) {

  bdims_summary <- bdims %>%
    summarize(N = n(), r = cor(hgt, wgt),
              mean_hgt = mean(hgt), mean_wgt = mean(wgt),
              sd_hgt = sd(hgt), sd_wgt = sd(wgt)) %>%
    mutate(true_slope = r * sd_wgt / sd_hgt, 
           true_intercept = mean_wgt - true_slope * mean_hgt)
  p <- ggplot(data = bdims, aes(x = hgt, y = wgt)) + 
    geom_point() + 
    geom_point(data = bdims_summary, 
               aes(x = mean_hgt, y = mean_wgt), 
               color = "red", size = 3)
  
  my_data <- bdims_summary %>%
    mutate(my_slope = my_slope, 
           my_intercept = mean_wgt - my_slope * mean_hgt)
  p + geom_abline(data = my_data, 
                  aes(intercept = my_intercept, slope = my_slope), color = "dodgerblue")
}

ggplot(data = bdims, aes(x = hgt, y = wgt)) + 
  geom_point()
```

```{r ex2, exercise = TRUE}
# Estimate optimal value of my_slope
 add_line(my_slope = ___)
```

<div id="ex2-hint">
**Hint:** Try a few values for the slope in the low single digits.
</div>

```{r ex2-solution}
# Estimate optimal value of my_slope
add_line(my_slope = 1)
```

## Understanding Linear Models

### Generic statistical model

Models are ubiquitous in statistics. In many cases, we assume that the value of our response variable is some function of our explanatory variable, plus some random noise. The latter term is important, and in a philosophical sense, is the signal of statistical thinking.

What we are saying here is that there is some mathematical function $f$, which can translate values of one variable into values of another, except that there is some randomness in the process. What often distinguishes statisticians from other quantitative researchers is the way that we try to model that random noise. 

$$
  response = f(explanatory) + noise
$$

### Generic linear model

For a linear regression model, we simply assume that $f$ takes the form of a linear function. Thus, our model describes the value of the response variable in terms of an intercept and a slope. 

$$
  response = intercept + slope \cdot explanatory + noise
$$

### Regression model

Mathematically, we usually call the intercept $\beta_0$, and the slope $\beta_1$. The noise term is often denoted $\epsilon$. In a regression model, we specify that the distribution of the noise is normal, with mean 0 and a fixed standard deviation. Again, understanding the specification of this noise term is crucial to thinking like a statistician. 

$$
  Y = \beta_0 + \beta_1 \cdot X + \epsilon \,, \qquad \epsilon \sim N(0, \sigma_\epsilon)
$$

### Fitted values

The part of the function aside from the noise term consists of a linear function that produces the fitted values. These are usually denoted with $\hat{Y}$. The difference between $Y$ and $\hat{Y}$, is that $Y$ is the actual observed values of the response, while $\hat{Y}$ is the expected values of the response based on the model. 

$$
  \hat{Y} = \hat{\beta}_0 + \hat{\beta}_1 \cdot X
$$

### Residuals

The difference between those two quantities are called residuals. The order of the terms here matters, so that observations that exceed their expected value of the response variable will have positive residuals, while those that don't reach their expected value will have negative residuals. 

The residuals are the realization of the noise term. While $\epsilon$ and $e$ play analogous roles in the regression equation, $\epsilon$ is an unknown, true quantity, while $e$ is a known, estimate of that quantity. 

$$
  e = Y - \hat{Y}
$$

### Fitting procedure

While we won't delve into the specifics of the least squares fitting procedure, we will outline what it does. Given $n$ observations of $(x,y)$-pairs, the procedure finds estimates for the intercept and slope that minimize the sum of the squared residuals. The estimated intercept and slope are usually denoted $\hat{\beta}_0$ and $\hat{\beta}_1$, respectively. Note that here again, the $\hat{\cdot}$ indicates an estimated---rather than a true---quantity. 

Mathematically, the steps of the process will be:

1. Given $n$ observations of pairs $(x_i, y_i)$
2. Find $\hat{\beta}_0, \hat{\beta}_1$ that minimize $\sum_{i=1}^n e_i^2$

### Least squares

Actually computing the fitted coefficients is usually a messy business for which the computer is much better suited than we are, so we won't discuss that here. You should know that the least squares fitting procedure is a well-understood, relatively straightforward, deterministic process that can be computed efficiently. It will always return a unique solution, except in rare, unrealistic cases. The residuals are guaranteed to sum to 0, and the point $(\bar{x}, \bar{y})$ is guaranteed to lie on the regression line. Given your understanding of correlation, it should not surprise you to learn that the regression slope and the correlation coefficient are closely-related. In fact, they are proportional to one another. 

To summarize, some aspects of the least squares are:
- Easy, deterministic, unique solution
- Residuals sum to zero
- Line must pass through $(\bar{x}, \bar{y})$
- Slope proportional to correlation
- Other criteria exist---just not in this tutorial

You should also be aware that other criteria---apart from least squares---exist, but we won't talk about them in this tutorial.

### Key concepts

It's worth reviewing some key concepts about regression models. First, the fitted values $\hat{Y}$ are the *expected* values given the corresponding value of $X$. It's not the case that $X$ will always translate into $\hat{Y}$; it's just that $\hat{Y}$ is our best guess for the true value of $Y$ given what we know about $X$. 

In the same way that the fitted intercept and slope are estimates of the true, unknown intercept and slope, the residuals are estimates of the true, unknown noise. 

Speaking of which, you may hear these noise terms referred to as "errors". We tend to shy away from this terminology, since it suggests that something is "wrong" with your model. But there is not necessarily anything wrong---it's just that physical processes are subject to some random variation. All we are hoping to do is accurately capture that random noise. 

To summarize:

- $\hat{Y}$ is *expected* value given corresponding $X$
- $\hat{\beta}$'s are estimates of true, unknown $\beta$'s
- Residuals ($e$'s) are estimates of true, unknown $\epsilon$'s
- "Error" may be misleading term---better: *noise*


You'll put your understanding of regression to use in these next exercises. 

### Regression model terminology

Consider a linear regression model of the form:

$$
    Y = \beta_0 + \beta_1 \cdot X + \epsilon \,, \text{ where } \epsilon \sim N(0, \sigma_{\epsilon}) \,.
$$

The slope coefficient is:

1. $Y$

2. $\beta_0$

3. $\beta_1$

4. $\epsilon$


```{r mc}
question("The slope coefficient is:",
  answer("1", message = "Not quite! 1 represents the response variable."),
  answer("2", message = "Try again! 2 represents the y-intercept."),
  answer("3", correct = TRUE, message = "The $X$ represents the explanatory variables and the $\beta_0$ represents the y-intercept."),
  answer("4", message = "Wrong! That represents the disturbance term."),
  allow_retry = TRUE
)
```


### Regression model output terminology

The fitted model for the poverty rate of U.S. counties as a function of high school graduation rate is:

$$
\widehat{poverty} = 64.594 - 0.591 \cdot hs_{grad}
$$


```{r mc1}
question("In Hampshire County in western Massachusetts, the high school graduation rate is 92.4%. These two facts imply that the poverty rate in Hampshire County is ___.
",
  answer("exactly 11.7%", message = "Incorrect, try again."),
  answer("exactly 10.0%", message = "Not quite!"),
  answer("expected to be about 10.0%", correct = TRUE, message = "Solve for poverty by plugging in 92.4 as hs_grad in the equation!"),
  answer("expected to be about 11.7%", message = "Did you plug 92.4 into the equation?"),
  allow_retry = TRUE
)
```


### Fitting a linear model "by hand"

Recall the simple linear regression model:
$$
    Y = b_0 + b_1 \cdot X
$$

Two facts enable you to compute the slope $b_1$ and intercept $b_0$ of a simple linear regression model from some basic summary statistics.

First, the slope can be defined as:

$$
    b_1 = r_{X,Y} \cdot \frac{s_Y}{s_X}
$$

where $r_{X,Y}$ represents the correlation (`cor()`) of $X$ and $Y$ and $s_X$ and $s_Y$ represent the standard deviation (`sd()`) of $X$ and $Y$, respectively.

Second, the point $(\bar{x}, \bar{y})$ is *always* on the least squares regression line, where $\bar{x}$ and $\bar{y}$ denote the average of $x$ and $y$, respectively.

The `bdims_summary` data frame contains all of the information you need to compute the slope and intercept of the least squares regression line for body weight ($Y$) as a function of height ($X$). You might need to do some algebra to solve for $b_0$!


- Print the `bdims_summary` data frame.
- Use `mutate()` to add the `slope` and `intercept` to the `bdims_summary` data frame.

```{r ex3-setup}
bdims_summary <- bdims %>%
  summarize(N = n(), r = cor(hgt, wgt),  
            mean_hgt = mean(hgt), sd_hgt = sd(hgt), 
            mean_wgt = mean(wgt), sd_wgt = sd(wgt))
```

```{r ex3, exercise = TRUE}
# Print bdims_summary


# Add slope and intercept
bdims_summary %>%
  mutate(slope = ___, 
         intercept = ___)
```

<div id="ex3-hint">
**Hint:** In order to calculate the `intercept`, we first calculate `slope`, then solve $Y = b_0 + b_1 \cdot X$ 
for $b_0$.
</div>

```{r ex3-solution}
# Print bdims_summary
bdims_summary

# Add slope and intercept
bdims_summary %>%
  mutate(slope = r * sd_wgt / sd_hgt, 
         intercept = mean_wgt - slope * mean_hgt)
```

<!--

## Regression vs. regression to the mean

### Heredity

In the late 19th century, Sir Francis Galton developed much of the theory surrounding correlation and regression. One of the more enduring concepts is "regression to the mean." Here we note that "regression to the mean" is a distinct concept from "linear regression", which is the focus of this tutorial. But "regression to the mean" is an important statistical concept in its own right, so we will explore it here so that you can appreciate the difference. 

As an example of Galton's "regression to the mean", consider analyzing  the heights of the children of NBA players.

As always, it's best to start with a question: Do tall men tend to beget tall sons? Do tall women tend to have tall daughters? You might consider this question in the context of NBA and WNBA players. Do you suspect that children of NBA and WNBA players are likely to be tall? 

### Galton's data

The answer turns out to be "yes, but not as tall as their fathers." This effect is what Galton deemed "regression to the mean". Another way of thinking about this is that unusual phenomena are likely to become more common over time.

In the plot, we see the relationship between the heights of a group of 465 men that Galton measured in the 1880s, and their fathers. The height of the son is on the $y$ axis, while the height of the corresponding father is on the $x$ axis. The diagonal indicates sons who are as tall as their fathers. The slope of the regression line---shown in blue---is more gentle. This reflects the observation that while tall fathers are still likely to have tall sons, the sons are not likely to be *as tall* as their fathers. Their height has been "regressed" towards the mean height. 

```{r}
Galton_men <- mosaicData::Galton %>%
 filter(sex == "M")
Galton_summary <- Galton_men %>%
 summarize(N = n(),
           mean_height = mean(height), sd_height = sd(height),
           mean_father = mean(father), sd_father = sd(father))
ggplot(data = Galton_men, aes(x = father, y = height)) +
 geom_point() +
# geom_rect(data = Galton_summary,
#           aes(xmin = mean_height, xmax = 100, ymin = mean_father, ymax = 100))
 geom_hline(data = Galton_summary, aes(yintercept = mean_height), lty = 3) +
 geom_vline(data = Galton_summary, aes(xintercept = mean_father), lty = 3) +
 geom_abline(slope = 1, intercept = 0, lty = 2) +
 geom_smooth(method = "lm", se = FALSE)
```

It may be tempting to think that regression to the mean implies that men are getting shorter, but this is not true. Even in this sample, the mean height of the sons was the same---to the nearest tenth of an inch---to those of the fathers. 

### Regression modeling

This tutorial is about simple linear regression *modeling*. Typically, regression models combine some explanatory variables into an estimate for a single, numerical response variable. We are pursuing least squares regression models, but there are any number of alternative regression model specifications. Additionally, there are models like "regression trees" that also estimate a single, numerical response, but do not use a linear framework as we do in this tutorial. 

"Regression": techniques for modeling a quantitative response can include a variety of models including:  
- Least squares  
- Weighted  
- Generalized  
- Nonparametric  
- Ridge  
- Bayesian  
- ...  
    
The next exercises will allow you to explore Galton's data on your own. 

### Regression to the mean

*Regression to the mean* is a concept attributed to Sir Francis Galton. The basic idea is that extreme random observations will tend to be less extreme upon a second trial. This is simply due to chance alone. While "regression to the mean" and "linear regression" are not the same thing, we will examine them together in this exercise.

One way to see the effects of regression to the mean is to compare the heights of parents to their children's heights. While it is true that tall mothers and fathers tend to have tall children, those children tend to be less tall than their parents, relative to average. That is, fathers who are 3 inches taller than the average father tend to have children who may be taller than average, but by less than 3 inches. 

The `Galton_men` and `Galton_women` datasets contain data originally collected by Galton himself in the 1880s on the heights of men and women, respectively, along with their parents' heights.

Compare the slope of the regression line to the slope of the diagonal line. What does this tell you?


- Create a scatterplot of the height of men as a function of their father's height. Add the simple linear regression line and a diagonal line (with slope equal to 1 and intercept equal to 0) to the plot.
- Create a scatterplot of the height of women as a function of their mother's height. Add the simple linear regression line and a diagonal line to the plot.

```{r ex4-setup}
Galton_men <- filter(Galton, sex == "M")
Galton_women <- filter(Galton, sex == "F")
```

```{r ex4, exercise = TRUE}
# Height of children vs. height of father
ggplot(data = ___, aes(x = ___, y = ___)) +
  geom_point() + 
  geom_abline(slope = ___, intercept = ___) + 
  geom_smooth(method = ___, se = FALSE)

# Height of children vs. height of mother
```

```{r ex4-hint}
"Use the scaffolding for the first scatterplot to generate your
second scatterplot."
```

```{r ex4-solution}
# Height of children vs. height of father
ggplot(data = Galton_men, aes(x = father, y = height)) +
  geom_point() + 
  geom_abline(slope = 1, intercept = 0) + 
  geom_smooth(method = "lm", se = FALSE)

# Height of children vs. height of mother
ggplot(data = Galton_women, aes(x = mother, y = height)) +
  geom_point() + 
  geom_abline(slope = 1, intercept = 0) + 
  geom_smooth(method = "lm", se = FALSE)
```


### "Regression" in the parlance of our time

In an opinion piece about nepotism [published in *The New York Times* in 2015](http://www.nytimes.com/2015/03/22/opinion/sunday/seth-stephens-davidowitz-just-how-nepotistic-are-we.html), economist Seth Stephens-Davidowitz wrote that:

> "Regression to the mean is so powerful that once-in-a-generation talent basically never sires once-in-a-generation talent. It explains why Michael Jordan’s sons were middling college basketball players and Jakob Dylan wrote two good songs. It is why there are no American parent-child pairs among Hall of Fame players in any major professional sports league."



```{r mc2}
question("The author is arguing that...",
  answer("Because of regression to the mean, an outstanding basketball player is likely to have sons that are as good at basketball as him.", message = "False, think about the previous example concerning the children of very tall parents."),
  answer("Because of regression to the mean, an outstanding basketball player is likely to have sons that are not good at basketball.", message = "Not quite!  Think about the previous example concerning the children of very tall parents."),
  answer("Because of regression to the mean, an outstanding basketball player is likely to have sons that are good at basketball, but not as good as him.", correct = TRUE, message ="That's right!  Basketball skill is likely similar to the previous example concerning the children of very tall parents."),
  answer("Linear regression is incapable of evaluating musical or athletic talent.", message = "Incorrect!"),
  allow_retry = TRUE
)
```

-->

## Congratulations!

You have successfully completed Lesson 3 in Tutorial 3: Introduction to Linear Models.  

What's next?

`r emo::ji("ledger")` [Full list of tutorials supporting OpenIntro::Introduction to Modern Statistics](https://openintrostat.github.io/ims-tutorials/)

`r emo::ji("spiral_notepad")` [Tutorial 3: Introduction to Linear Models Data](https://openintrostat.github.io/ims-tutorials/03-introduction-to-linear-models/)

`r emo::ji("one")` [Tutorial 3 - Lesson 1: Visualizing two variables](https://openintro.shinyapps.io/ims-03-introduction-to-linear-models-01/)

`r emo::ji("two")` [Tutorial 3 - Lesson 2: Correlation](https://openintro.shinyapps.io/ims-03-introduction-to-linear-models-02/)

`r emo::ji("three")` [Tutorial 3 - Lesson 3: Simple linear regression](https://openintro.shinyapps.io/ims-03-introduction-to-linear-models-03/)

`r emo::ji("four")` [Tutorial 3 - Lesson 4: Interpreting regression models](https://openintro.shinyapps.io/ims-03-introduction-to-linear-models-04/)

`r emo::ji("five")` [Tutorial 3 - Lesson 5: Model fit](https://openintro.shinyapps.io/ims-03-introduction-to-linear-models-05/)

`r emo::ji("open_book")` [Learn more at Introduction to Modern Statistics](http://openintro-ims.netlify.app/)
